---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# ESTER: Efficient Sequential Testing with Evidence Ratios

[![CRAN status](http://www.r-pkg.org/badges/version/ESTER)](https://cran.r-project.org/package=ESTER) [![Build Status](https://travis-ci.org/lnalborczyk/ESTER.svg?branch=master)](https://travis-ci.org/lnalborczyk/ESTER)

`ESTER` offers an implementation of sequential testing that uses evidence ratios computed from the Akaike weights and based on either the Akaike Information Criterion (AIC) or the Bayesian Information Criterion (BIC).

## Installation

You can install ESTER from github with:

```{r gh-installation, eval = FALSE}
if(!require(devtools)){install.packages("devtools")}
devtools::install_github("lnalborczyk/ESTER")
```

## Different questions

1. **Simulation**. Given an expected effect size and sample size, what ER evolution should I reasonnably expect ?

2. **Observed data**. When to stop recruiting participants ?

## Simulation

This first function runs a simulated study in which we compare two independant groups, for various effect size and sample size. The `nmin` argument serves to specify from which participant we want to start doing sequential testing (usually we recommand to avoid `nmin` < 20).

```{r}
library(ESTER)
ER <- simER(cohensd = 0.6, nmin = 20, n = 100, ic = aic, plot = TRUE)
```

We also can study the distribution of evidence ratios for `nSims` simulations ran with the previous function using `distER`.

```{r}
ER <- distER(cohensd = 0.6, nmin = 20, n = 100, ic = aic, nsims = 100)
```

## Observed data

On the other hand (and perhaps more interestingly), `ESTER` can be used to do sequential testing on your own data. You can study the evolution of sequentials ERs using the `seqER` function.

```{r, results = "hide"}
data(mtcars)
mod1 <- lm(mpg ~ cyl, mtcars)
mod2 <- lm(mpg ~ cyl + disp, mtcars)
plot(seqER(ic = bic, mod1, mod2, nmin = 10) )
```

In addition, `seqERboot` allows you to study the behavior of sequential ERs computed on your own data, along with sequential ERs computed on bootstrapped samples from your data.

```{r, results = "hide"}
data(mtcars)
mod1 <- lm(mpg ~ cyl, mtcars)
mod2 <- lm(mpg ~ cyl + disp, mtcars)
plot(seqERboot(ic = bic, mod1, mod2, nmin = 10, order_nb = 20, replace = TRUE) )
```
