---
title: "Efficient Sequential Testing with Evidence Ratios"
author: "Ladislas Nalborczyk"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
---

This document demonstrates how to do sequential analyses for studies which use AICc-based Evidence Ratios (ER) as index of evidence.

## Installation

The `ESTER` package is not on CRAN yet, but you can install the development version from Github:

```{r, eval = FALSE}

library(devtools)
devtools::install_github("lnalborczyk/ESTER")

```

## Different questions

1. Simulation. Given an expected effect size and sample size, what ER evolution should I reasonnably expect ?
2. Observed data. Does the ER computed on my observed sample reached a stable region ? When stop to recruit participants ?

## 1. Simulation

```{r eval = FALSE}

ER <- simER(cohensd = 0.4, n = 60, nmin = 20, plot = TRUE)

```

## 2. Observed data.

```{r eval = FALSE}

library(lme4)
data <- sleepstudy
mod1 <- lm(Reaction ~ 1, data)
mod2 <- lm(Reaction ~ Days, data)
seqERboot(mod1, mod2, samplecol = "Subject", order_nb = 10, nmin = 10, replace = FALSE)

```

## Development

The *ESTER* package is also available in a development version at the GitHub development page: [https://github.com/lnalborczyk/ESTER](https://github.com/lnalborczyk/ESTER).
